---
title: "CO2 Emissions Responsibility Redesign"
author: "CSC3107 Team Project"
date: "2026-06-16"
format:
  html:
    toc: true
    embed-resources: true
execute:
  echo: true
  warning: false
  message: false
---

## Design focus

This redesign is based on the Reuters Graphics article "Who is the biggest polluter of carbon dioxide?" The aim is not to make a general climate-change poster. The aim is to show that national responsibility changes depending on how CO2 emissions are measured.

The working design question is:

> Who appears most responsible for CO2 emissions under annual, cumulative, per-capita, and consumption-based accounting?

## Data import and cleaning

```{r}
library(tidyverse)
library(scales)

raw_path <- file.path("data", "raw", "owid-co2-data.csv")
cleaned_dir <- file.path("data", "cleaned")
figures_dir <- "figures"

dir.create(cleaned_dir, recursive = TRUE, showWarnings = FALSE)
dir.create(figures_dir, recursive = TRUE, showWarnings = FALSE)

stopifnot(file.exists(raw_path))

owid <- read_csv(raw_path, show_col_types = FALSE)

available_countries <- unique(owid$country)
eu_label <- intersect(
  c("European Union (27)", "EU-27", "European Union"),
  available_countries
)[1]

comparison_countries <- c(
  "China",
  "United States",
  "India",
  eu_label,
  "Russia",
  "Japan",
  "Germany",
  "United Kingdom",
  "Singapore"
)

comparison_countries <- comparison_countries[!is.na(comparison_countries)]

missing_countries <- setdiff(comparison_countries, available_countries)
if (length(missing_countries) > 0) {
  stop("Missing expected country labels: ", paste(missing_countries, collapse = ", "))
}

metric_columns <- c(
  "co2",
  "cumulative_co2",
  "co2_per_capita",
  "consumption_co2_per_capita"
)

analysis_year <- owid |>
  filter(country %in% comparison_countries) |>
  filter(if_all(all_of(metric_columns), ~ !is.na(.x))) |>
  count(year, name = "complete_countries") |>
  filter(complete_countries == length(comparison_countries)) |>
  summarise(year = max(year)) |>
  pull(year)

if (length(analysis_year) == 0 || is.na(analysis_year)) {
  stop("No single year has all required metrics for the selected countries.")
}

cleaned <- owid |>
  filter(country %in% comparison_countries, year == analysis_year) |>
  select(
    country,
    year,
    iso_code,
    population,
    co2,
    co2_per_capita,
    cumulative_co2,
    consumption_co2,
    consumption_co2_per_capita,
    share_global_co2,
    share_global_cumulative_co2
  ) |>
  arrange(match(country, comparison_countries))

write_csv(cleaned, file.path(cleaned_dir, "co2_selected_countries.csv"))

analysis_year
```

The selected analysis year is `r analysis_year`, because it is the latest year where all selected comparison countries have complete values for the four required metrics.

## Transform for visualization

```{r}
metric_levels <- c(
  "Annual territorial CO2",
  "Cumulative territorial CO2",
  "Territorial CO2 per capita",
  "Consumption CO2 per capita"
)

plot_data <- cleaned |>
  transmute(
    country,
    `Annual territorial CO2` = co2 / 1000,
    `Cumulative territorial CO2` = cumulative_co2 / 1000,
    `Territorial CO2 per capita` = co2_per_capita,
    `Consumption CO2 per capita` = consumption_co2_per_capita
  ) |>
  pivot_longer(
    cols = -country,
    names_to = "metric",
    values_to = "value"
  ) |>
  mutate(
    metric = factor(metric, levels = metric_levels),
    unit = if_else(str_detect(as.character(metric), "per capita"), "tonnes/person", "Gt CO2"),
    value_label = case_when(
      unit == "Gt CO2" ~ number(value, accuracy = 0.1),
      TRUE ~ number(value, accuracy = 0.1)
    )
  ) |>
  group_by(metric) |>
  mutate(rank = min_rank(desc(value))) |>
  ungroup()

write_csv(plot_data, file.path(cleaned_dir, "co2_metric_comparison_long.csv"))

plot_data |>
  arrange(metric, rank) |>
  select(metric, rank, country, value, unit)
```

## Draft redesign

```{r}
country_palette <- c(
  "China" = "#c7512f",
  "United States" = "#2f6f9f",
  "India" = "#7a5ea8",
  "European Union (27)" = "#4d8c57",
  "EU-27" = "#4d8c57",
  "European Union" = "#4d8c57",
  "Russia" = "#8f5d46",
  "Japan" = "#d08c2f",
  "Germany" = "#6b6b6b",
  "United Kingdom" = "#3f7f7a",
  "Singapore" = "#b64d7a"
)

country_levels <- rev(comparison_countries)

small_multiple_chart <- ggplot(
  plot_data,
  aes(
    x = value,
    y = factor(country, levels = country_levels),
    fill = country
  )
) +
  geom_col(width = 0.68) +
  geom_text(
    aes(label = value_label),
    hjust = -0.08,
    size = 3.1,
    color = "#222222"
  ) +
  facet_wrap(~ metric, scales = "free_x", ncol = 2) +
  scale_fill_manual(values = country_palette, guide = "none") +
  scale_x_continuous(
    labels = label_number(accuracy = 1),
    expand = expansion(mult = c(0, 0.22))
  ) +
  labs(
    title = "The biggest polluter changes depending on the metric",
    subtitle = paste0(
      "Selected countries, ",
      analysis_year,
      ". Annual and cumulative emissions are shown in gigatonnes CO2; per-capita measures are tonnes per person."
    ),
    x = NULL,
    y = NULL,
    caption = "Data: Our World in Data CO2 and Greenhouse Gas Emissions dataset. Reuters graphic source also cites Our World in Data."
  ) +
  theme_minimal(base_size = 11) +
  theme(
    plot.title = element_text(face = "bold", size = 16),
    plot.subtitle = element_text(size = 10, color = "#444444"),
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank(),
    strip.text = element_text(face = "bold", size = 11),
    axis.text.y = element_text(size = 9),
    axis.text.x = element_text(size = 8),
    plot.caption = element_text(size = 8, color = "#555555"),
    panel.spacing = unit(1.2, "lines")
  ) +
  coord_cartesian(clip = "off")

small_multiple_chart

ggsave(
  filename = file.path(figures_dir, "co2_metric_small_multiples.png"),
  plot = small_multiple_chart,
  width = 12,
  height = 8,
  dpi = 300
)
```

## Starter interpretation

The first chart draft is meant to support a narrow poster argument:

- Annual territorial emissions emphasize current industrial scale.
- Cumulative territorial emissions emphasize historical responsibility.
- Territorial per-capita emissions emphasize average domestic emissions intensity.
- Consumption-based per-capita emissions shift attention toward demand and trade.

The poster should explain that none of these metrics is a neutral single answer. Each one answers a different policy question.

## Alternative Chart Implementations

This section explores other chart types that could improve upon the original Reuters graphic, focusing on rank shifts across different perspectives.

### Slope Chart (Paired)

The paired slope chart simplifies the comparison by breaking it down into two side-by-side panels. It effectively communicates the change in ranking between two metrics at a time.

```{r}
# Prepare data for paired slope chart
slope_rank_data <- plot_data |>
  select(country, metric, rank) |>
  mutate(
    metric_short = case_when(
      metric == "Cumulative territorial CO2" ~ "Cumulative",
      metric == "Annual territorial CO2" ~ "Annual",
      metric == "Territorial CO2 per capita" ~ "Per capita",
      metric == "Consumption CO2 per capita" ~ "Consumption per cap"
    )
  )

# Abbreviate long country names for slope chart labels
slope_rank_data <- slope_rank_data |>
  mutate(
    country_short = case_when(
      country == "European Union (27)" ~ "EU-27",
      country == "United Kingdom" ~ "UK",
      TRUE ~ country
    )
  )

# Create two paired comparisons 
pair1 <- slope_rank_data |>
  filter(metric_short %in% c("Cumulative", "Annual")) |>
  mutate(
    panel = "Cumulative \u2192 Annual",
    x_pos = if_else(metric_short == "Cumulative", 1, 4)
  )

pair2 <- slope_rank_data |>
  filter(metric_short %in% c("Annual", "Per capita")) |>
  mutate(
    panel = "Annual \u2192 Per capita",
    x_pos = if_else(metric_short == "Annual", 1, 4)
  )

slope_plot_data <- bind_rows(pair1, pair2) |>
  mutate(
    panel = factor(panel, levels = c("Cumulative \u2192 Annual", "Annual \u2192 Per capita")),
    label_text = paste0("#", rank, " ", country_short),
    is_left = x_pos == 1,
    hjust = if_else(is_left, 1, 0),
    label_x = if_else(is_left, x_pos - 0.2, x_pos + 0.2)
  )

# Axis column headers
axis_labels <- data.frame(
  panel = factor(
    c("Cumulative \u2192 Annual", "Cumulative \u2192 Annual", 
      "Annual \u2192 Per capita", "Annual \u2192 Per capita"),
    levels = c("Cumulative \u2192 Annual", "Annual \u2192 Per capita")
  ),
  x_pos = c(1, 4, 1, 4),
  y_pos = 0.2,
  label = c("Cumulative", "Annual", "Annual", "Per capita"),
  hjust = c(1, 0, 1, 0)
)

slope_chart <- ggplot(slope_plot_data, aes(y = rank, group = country, color = country)) +
  geom_line(aes(x = x_pos), linewidth = 1.2) +
  geom_point(aes(x = x_pos), size = 3) +
  geom_text(
    aes(x = label_x, label = label_text, hjust = hjust),
    size = 3.5,
    fontface = "bold"
  ) +
  geom_text(
    data = axis_labels,
    aes(x = x_pos, y = y_pos, label = label, hjust = hjust),
    inherit.aes = FALSE,
    fontface = "bold",
    size = 4,
    color = "#333333"
  ) +
  facet_wrap(~ panel, scales = "free_x") +
  scale_y_reverse(limits = c(max(slope_plot_data$rank) + 0.5, 0)) +
  scale_x_continuous(
    limits = c(-2, 7),
    expand = c(0, 0)
  ) +
  scale_color_manual(values = country_palette, guide = "none") +
  labs(
    title = "Slope chart (paired)",
    subtitle = "Cumulative vs Annual vs Per Capita Emissions",
    x = NULL, y = NULL
  ) +
  theme_minimal(base_size = 11) +
  theme(
    plot.title = element_text(face = "bold", size = 16),
    plot.subtitle = element_text(size = 10, color = "#444444"),
    panel.grid = element_blank(),
    axis.text.y = element_blank(),
    axis.text.x = element_blank(),
    strip.text = element_text(face = "bold", size = 12),
    panel.spacing = unit(4, "lines"),
    plot.margin = margin(10, 30, 10, 30)
  ) +
  coord_cartesian(clip = "off")

slope_chart

ggsave(
  filename = file.path(figures_dir, "co2_slope_chart.png"),
  plot = slope_chart,
  width = 14,
  height = 7,
  dpi = 300
)
```

## Next design tasks

- Decide whether to keep the EU aggregate or remove it to avoid mixing countries and regions.
- Add concise annotations beside each panel.
- Add a small supporting rank-change chart only if the poster has enough space.
- Place the original Reuters graphic beside a short critique of the Sankey and rotated-label design.
