---
title: "CO2 Redesign: Chart Alternatives"
subtitle: "Testing bump chart and connected dot plot against the current faceted bar chart"
format:
  html:
    toc: true
    embed-resources: true
execute:
  echo: true
  warning: false
  message: false
---

## Why test alternatives

The current main chart (`co2_redesign.qmd`) uses a faceted bar chart: one ranked
bar panel per metric. It is easy to read exact values, but it breaks the thing
the original Reuters ribbon chart did well — letting a reader watch one
country's position shift as the metric changes. This page builds two
alternatives that keep that "story" while fixing the original's real problems
(uncontrolled flow widths, dense crossings, unreadable rotated labels for small
countries).

```{r}
library(tidyverse)

plot_data <- read_csv("../data/cleaned/co2_metric_comparison_long.csv", show_col_types = FALSE)

metric_levels <- c(
  "Annual territorial CO2",
  "Cumulative territorial CO2",
  "Territorial CO2 per capita",
  "Consumption CO2 per capita"
)

plot_data <- plot_data |>
  mutate(metric = factor(metric, levels = metric_levels))

country_palette <- c(
  "China" = "#E69F00",
  "United States" = "#56B4E9",
  "India" = "#009E73",
  "European Union (27)" = "#F0E442",
  "Russia" = "#0072B2",
  "Japan" = "#D55E00",
  "Germany" = "#CC79A7",
  "United Kingdom" = "#999999",
  "Singapore" = "#882255"
)

# Countries to highlight in color; everyone else fades to gray context lines.
highlight_countries <- c("China", "United States", "India", "Singapore")

plot_data <- plot_data |>
  mutate(
    highlight = country %in% highlight_countries,
    line_color = if_else(highlight, country, "Other")
  )

highlight_palette <- c(country_palette[highlight_countries], "Other" = "#cccccc")
```

## Option 1: Bump chart

Rank (not raw value) on the y-axis, metric on the x-axis, one line per country.
This is the most direct structural replacement for the original ribbon
chart's "rank-flow" story.

```{r}
bump_chart <- ggplot(
  plot_data,
  aes(x = metric, y = rank, group = country, color = line_color)
) +
  geom_line(aes(linewidth = highlight), lineend = "round") +
  geom_point(aes(size = highlight)) +
  geom_text(
    data = filter(plot_data, metric == first(metric_levels), highlight),
    aes(label = country),
    hjust = 1.15, size = 3.2, fontface = "bold"
  ) +
  geom_text(
    data = filter(plot_data, metric == last(metric_levels), highlight),
    aes(label = country),
    hjust = -0.15, size = 3.2, fontface = "bold"
  ) +
  scale_y_reverse(breaks = 1:9) +
  scale_color_manual(values = highlight_palette, guide = "none") +
  scale_linewidth_manual(values = c(`TRUE` = 1.1, `FALSE` = 0.5), guide = "none") +
  scale_size_manual(values = c(`TRUE` = 3, `FALSE` = 1.5), guide = "none") +
  labs(
    title = "Rank shifts depending on how you measure CO2 responsibility",
    subtitle = "Highlighted: China, United States, India, Singapore. Other countries shown as context.",
    x = NULL, y = "Rank (1 = highest)"
  ) +
  theme_minimal(base_size = 11) +
  theme(
    panel.grid.minor = element_blank(),
    plot.margin = margin(10, 60, 10, 60)
  ) +
  coord_cartesian(clip = "off")

bump_chart
ggsave("bump_chart.png", bump_chart, width = 10, height = 6, dpi = 300)
```

## Option 2: Connected dot plot

Countries as rows, one dot per metric, rank shown as the label inside each
dot, dots in a row connected by a thin line. Less dramatic than the bump
chart's crossing lines, but easier to scan per country.

```{r}
country_order <- plot_data |>
  filter(metric == "Annual territorial CO2") |>
  arrange(rank) |>
  pull(country)

dot_data <- plot_data |>
  mutate(country = factor(country, levels = rev(country_order)))

connected_dot_plot <- ggplot(dot_data, aes(x = metric, y = country, group = country)) +
  geom_line(color = "#999999", linewidth = 0.6) +
  geom_point(aes(fill = country), shape = 21, size = 8, color = "white", stroke = 0.6) +
  geom_text(aes(label = rank), size = 3, fontface = "bold", color = "#222222") +
  scale_fill_manual(values = country_palette, guide = "none") +
  labs(
    title = "Each country's rank across four CO2 accounting metrics",
    x = NULL, y = NULL
  ) +
  theme_minimal(base_size = 11) +
  theme(
    panel.grid.major.y = element_blank(),
    axis.text.x = element_text(angle = 20, hjust = 1)
  )

connected_dot_plot
ggsave("connected_dot_plot.png", connected_dot_plot, width = 10, height = 6, dpi = 300)
```

## Discussion notes for the team

- **Bump chart**: closest structural match to the brief ("translate from
  perspective to perspective easily") and to what the professor described.
  With 9 countries, full-color lines get crowded — used highlight + gray
  context here. Decide as a team whether 4 highlighted countries is the
  right number, or whether it should track the same comparison set as the
  main chart.
- **Connected dot plot**: compact, shows the exact rank number without
  needing to read position on a y-axis, but loses the "crossing lines = the
  story" visual impact that sold the original chart's narrative.
- **Faceted bar chart (existing, not reproduced here)**: still the best
  option if precise values matter more than rank-flow storytelling — worth
  keeping as a secondary panel either way, even if the primary chart changes.
